Exploratory Data Analysis for College Scorecard Dataset

### 2017-02-22 Draft-xuezhou-sjb.Rmd
update family income

#Table of Contents
* 1.[Introduction](#Introduction)
* 2.[Goal](#Goal)
* 3.[Dataset Preparation](#dataprep)
* 4.[Objective](#objective)
* 5.[Summary of Variables](#variablesummary)
* 5.1 [Individual Variable Summary](#5.1)
  * 5.1.1 [School](#5.1.1)
    * 5.1.1.1 [State](#5.1.1.1)
    * 5.1.1.2 [Predominant Degree](#5.1.1.2)
    * 5.1.1.3 [Control of Institution](#5.1.1.3)
  * 5.1.2 Student Body
    * 5.1.2.1 [Percentage First-generation Students](#5.1.2.1)
    * 5.1.2.2 [Number of Applications](#5.1.2.2)
  * 5.1.3 Cost 
    * 5.1.3.1 [Average Net Price for Title IV Institutions (Both Public and Private Institutions)](#5.1.3.1)
    * 5.1.3.2 [Percentage of Students Who are Financially Independent](#5.1.3.2)
  * 5.1.4 Financial Aid
    * 5.1.4.1 [Percentage of Undergraduates Who Receive a Pell Grant](#5.1.4.1)
    * 5.1.4.2 [Percent of all Federal Undergraduate Students Receiving a Federal Student Loan](#5.1.4.2l)
    * 5.1.4.3 [The Median Debt for Students Who Have Completed](#5.1.4.3)
    * 5.1.4.4 [The Median Debt for Students Who Have not Completed](#5.1.4.4)
  * 5.1.5 Earning
    * 5.1.5.1 [Median Family Income in real 2015 Dollars](#5.1.5.1)
    * 5.1.5.2 [Earning after graduation](#5.1.5.2)
  * 5.1.6 Repayment
    * 5.1.6.1 [3-year Repayment rate](#5.1.6.1)
    * 5.1.6.2 [3-year Default rate](#5.1.6.2)
* 6.Relationship between variables
 * 6.1 [The Relationship between Average Net Cost for Title IV institutions and State](#6.1)
 * 6.2 [The Relationship between Median Debt and Average Net Price for Title IV institutions](#6.2)
 * 6.3 [The Relationship between Percentage of Pell Grant and College Type (CONTROL)](#6.3)
 * 6.4 [The Relationship between Percentage First-generation Students and States](#6.4)
 * 6.5 [The Relationship between Median Family Income and Percentage of Pell Grant](#6.5)
 * 6.6 [The Relationship between the Percentage of Financially-independent Students and College Type (Control)](#6.6)
 * 6.7 [The Relationship between Percentage of Students Who Receive the Federal Loans and College Type (Control).](#6.7)
 * 6.8 [The Relationsip between Median Debts and Post-Graduation Earnings](#6.8)
 * 6.9 [The Relationship between the Default Rate of Loan Repayment and Post-graduation Earnings ](#6.9)
 * 6.10 [The Relationship between Earning 6 Years from Entry and Degree Type](#6.10)
 * 6.11 [The Relationship among Predominant Degree, College Type(Control) and Earning after 6 year of Entry](#6.11)
 * 6.12 [The Relationship between Predominant Degree and College Type (Control)](#6.12)
 * 6.13 [The Relationship between College Type (Control) and Median earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
* 7. [Findings and Conclusion](#7)

@@

| Variable Name    | Data Type |  Unit  |  Description                                                          |    Category in Documentation       |
| --------------   | ----------|--------|-----------------------------------------------------------------------| --------------------------------|
| NPT4_COMBINE    | Numeric   | $         |average net price for TITLE IV institutions of both public and private.|  Cost                       |
| DEP_STAT_PCT_IND | Numeric  | %         |Percentage of students who are financially independent                |     Cost                        |
| PAR_ED_PCT_1STGEN| Numeric  | %         |Percentage first-generation students                                  |      Sudent Body                |
| MD_FAMINC        | Numeric  | $         |Median family income in real 2015 dollars                             |       Earning                   |
| PCTPELL          | Numeric  | %         |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
| PCTFLOAN         | Numeric  | %         |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
| GRAD_DEBT_MDN    | Numeric  | $         |The median debt for students who have completed                       |   Financial Aid                 |
| WDRAW_DEBT_MDN   | Numeric  | $         |The median debt for sutdents who have not completed                   |Financial Aid |
| MD_EARN_WNE_P6   | Numeric  | $         |Median income six years after entry                                   |     earning   |
| MD_EARN_WNE_P10  | Numeric  | $         |Median income ten years after entry                                   |     earning  |
| RPY_3YR_RT_SUPP  | Numeric  | %         |Repayment rate 3 years after entering repayment                       |Repayment   |
| CDR3             | Numeric  | %         |Loan default rate 3 years afer entering repayment                     |Repayment        |
| PREDDEG_factor    | Factor  | N/A          |Predominant degree offered                                            |School       |
| Control_factor    | Factor  | N/A          |Control of institution                                                |School     |
| STABBR_factor  | Factor     | N/A       |State                                                                 |School     |
| APPL_SCH_PCT_GE2 | Numeric  | %         |Percentage of students submitting 2 or more than 2 applications       | Student        |
| UNITID | Numeric | N/A      | Unique ID for the institutions | Root  |
| INSTNM | Factor |  N/A      | name for the institutions       | Institution |

@@

@@

###5.1.5 Earning<a id="5.1.5"></a> 
####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts. We are using the code below to summarize this variable(min, max, mean, median and standard deviation).

####5.1.5.2 Earning after graduation<a id="5.1.5.2"></a>

@@

From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data shows a downhill pattern from left to right; As the MD_FAMINC increases, the value of PCTPELL decreases. It makes sense in the real world; for institutions with low percentage of students who receive the Pell Grant, the financial aid like Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, since these collges students tend to have a higher median family income.

@@

##6.8 The Relationsip between Median Debts and Post-Graduation Earnings<a id="6.8"></a> 
Next, we analyze the relationship between the amount of the debt and post-graduation earnings. There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study. We decided to use debt for students who have not completed their study in this problem, since it makes more sense for the purpose of prediction.

@@

It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, thus these college students will tend to have a high median family income after they graduate.

>>>>

#Table of Contents
* 1.[Introduction](#Introduction)
* 2.[Goal](#Goal)
* 3.[Dataset Preparation](#dataprep)
* 4.[Objective](#objective)
* 5.[Summary of Variables](#variablesummary)
* 5.1 [Individual Variable Summary](#5.1)
  * 5.1.1 [School](#5.1.1)
    * 5.1.1.1 [State](#5.1.1.1)
    * 5.1.1.2 [Predominant Degree](#5.1.1.2)
    * 5.1.1.3 [Control of Institution](#5.1.1.3)
  * 5.1.2 Student Body
    * 5.1.2.1 [Percentage First-generation Students](#5.1.2.1)
    * 5.1.2.2 [Number of Applications](#5.1.2.2)
    * 5.1.2.3 [Median Family Income in Real 2015 Dollars](#5.1.2.3)
  * 5.1.3 Cost 
    * 5.1.3.1 [Average Net Price for Title IV Institutions (Both Public and Private Institutions)](#5.1.3.1)
    * 5.1.3.2 [Percentage of Students Who are Financially Independent](#5.1.3.2)
  * 5.1.4 Financial Aid
    * 5.1.4.1 [Percentage of Undergraduates Who Receive a Pell Grant](#5.1.4.1)
    * 5.1.4.2 [Percent of all Federal Undergraduate Students Receiving a Federal Student Loan](#5.1.4.2l)
    * 5.1.4.3 [The Median Debt for Students Who Have Completed](#5.1.4.3)
    * 5.1.4.4 [The Median Debt for Students Who Have not Completed](#5.1.4.4)
  * 5.1.5 Earning
    * 5.1.5.1 [Earning after graduation](#5.1.5.1)
  * 5.1.6 Repayment
    * 5.1.6.1 [3-year Repayment rate](#5.1.6.1)
    * 5.1.6.2 [3-year Default rate](#5.1.6.2)
* 6.Relationship between variables
 * 6.1 [The Relationship between Average Net Cost for Title IV institutions and State](#6.1)
 * 6.2 [The Relationship between Median Debt and Average Net Price for Title IV institutions](#6.2)
 * 6.3 [The Relationship between Percentage of Pell Grant and College Type (CONTROL)](#6.3)
 * 6.4 [The Relationship between Percentage First-generation Students and States](#6.4)
 * 6.5 [The Relationship between Median Family Income and Percentage of Pell Grant](#6.5)
 * 6.6 [The Relationship between the Percentage of Financially-independent Students and College Type (Control)](#6.6)
 * 6.7 [The Relationship between Percentage of Students Who Receive the Federal Loans and College Type (Control).](#6.7)
 * 6.8 [The Relationsip between Median Debts and Post-Graduation Earnings](#6.8)
 * 6.9 [The Relationship between the Default Rate of Loan Repayment and Post-graduation Earnings ](#6.9)
 * 6.10 [The Relationship between Earning 6 Years from Entry and Degree Type](#6.10)
 * 6.11 [The Relationship among Predominant Degree, College Type(Control) and Earning after 6 year of Entry](#6.11)
 * 6.12 [The Relationship between Predominant Degree and College Type (Control)](#6.12)
 * 6.13 [The Relationship between College Type (Control) and Median Earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
* 7. [Findings and Conclusion](#7)

@@

| Variable Name    | Data Type |  Unit  |  Description                                                          |    Category in Documentation       |
| --------------   | ----------|--------|-----------------------------------------------------------------------| --------------------------------|
| NPT4_COMBINE    | Numeric   | $         |average net price for TITLE IV institutions of both public and private.|  Cost                       |
| DEP_STAT_PCT_IND | Numeric  | %         |Percentage of students who are financially independent                |     Cost                        |
| PAR_ED_PCT_1STGEN| Numeric  | %         |Percentage first-generation students                                  |      Sudent Body                |
| MD_FAMINC        | Numeric  | $         |Median family income in real 2015 dollars                             |    Student Body                   |
| PCTPELL          | Numeric  | %         |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
| PCTFLOAN         | Numeric  | %         |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
| GRAD_DEBT_MDN    | Numeric  | $         |The median debt for students who have completed                       |   Financial Aid                 |
| WDRAW_DEBT_MDN   | Numeric  | $         |The median debt for sutdents who have not completed                   |Financial Aid |
| MD_EARN_WNE_P6   | Numeric  | $         |Median income six years after entry                                   |     earning   |
| MD_EARN_WNE_P10  | Numeric  | $         |Median income ten years after entry                                   |     earning  |
| RPY_3YR_RT_SUPP  | Numeric  | %         |Repayment rate 3 years after entering repayment                       |Repayment   |
| CDR3             | Numeric  | %         |Loan default rate 3 years afer entering repayment                     |Repayment        |
| PREDDEG_factor    | Factor  | N/A          |Predominant degree offered                                            |School       |
| Control_factor    | Factor  | N/A          |Control of institution                                                |School     |
| STABBR_factor  | Factor     | N/A       |State                                                                 |School     |
| APPL_SCH_PCT_GE2 | Numeric  | %         |Percentage of students submitting 2 or more than 2 applications       | Student        |
| UNITID | Numeric | N/A      | Unique ID for the institutions | Root  |
| INSTNM | Factor |  N/A      | name for the institutions       | Institution |

@@

####5.1.2.3 Median family income in real 2015 dollars<a id="5.1.2.3"></a>
This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts. We are using the code below to summarize this variable(min, max, mean, median and standard deviation).

@@

###5.1.5 Earning<a id="5.1.5"></a>

####5.1.5.1 Earning after graduation<a id="5.1.5.1"></a>

@@

From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data shows a downhill pattern from left to right; As the MD_FAMINC increases, the value of PCTPELL decreases. It makes sense in the real world; for institutions with low percentage of students who receive the Pell Grant, the financial aid like Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, as these college students may have enough median family income to cover their costs.

@@

##6.8 The Relationship between Median Debts and Post-Graduation Earnings<a id="6.8"></a> 
Next, we analyze the relationship between the amount of the debt and post-graduation earnings. There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study. We decided to use debt for students who have not completed their study in this problem, since it makes more sense for the purpose of prediction.

@@

It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, since these college students may have sufficient family income to support their education.
